---
title: "Exercise 1.13"
output: pdf_document
date: "2025-08-28"
---

Prove that Fib($n$) is the closest integer to $\phi^n/\sqrt{5}$, where $\phi = (1 + \sqrt{5})/2$. Hint: Let $\psi = (1 - \sqrt{5})/2$. Use induction and the definition of the Fibonacci numbers to prove that Fib($n$) = $(\phi^n - \psi^n)/\sqrt{5}$.\

$$
\text{Fib}(n) = \begin{cases}
  0, & \text{if } n = 0, \\
  1, & \text{if } n = 1, \\
  \text{Fib}(n - 1) + \text{Fib}(n - 2) & \text{otherwise.}
\end{cases}
$$

**Proposition.** Fib($n$) = $(\phi^n - \psi^n)/\sqrt{5}$.\

*Proof*. We will prove this with mathematical induction.\

(1) If $n = 0$, Fib(0) $= 0$. $\phi^0 = (\frac{(1 + \sqrt{5})}{2})^0 = 1$ and $\psi^1 =(\frac{(1 - \sqrt{5})}{2})^0 = 1$.\

$\frac{\phi^1 - \psi^1}{5} = \frac{1 - 1}{5} = \frac{0}{5} = 0$\

(2) Say $k \in \mathbb{N}$. Using direct proof, we will show that Fib($k$) = $(\phi^k - \psi^k)/\sqrt{5}$ implies that Fib($k + 1$) = $\frac{\phi^{k + 1} - \psi^{k + 1}}{\sqrt{5}}$\
Using the definition of the Fibonacci numbers, Fib($k + 1$) = Fib($(k + 1) - 1$) + Fib($(k + 1) - 2$) = Fib($k$) + Fib($k - 1$).\
Fib($k + 1$) = $\frac{\phi^k - \psi^k}{\sqrt{5}} + \frac{\phi^{k - 1} - \psi^{k - 1}}{\sqrt{5}} = \frac{\phi^k - \psi^k + \phi^{k - 1} - \psi^{k - 1}}{\sqrt{5}} = \frac{\phi^{k-1}(1 + \phi) - \psi^{k - 1}(1 + \psi)}{\sqrt{5}}$\
Remember, $\phi^2 = \phi + 1$\
We can also show that $\psi^2 = \psi + 1$\
$\psi^2 = (\frac{1 - \sqrt{5}}{2})^2 = \frac{(1 - \sqrt{5})(1 - \sqrt{5})}{4} = \frac{1 - 2\sqrt{5} + 5}{4} = \frac{6 - 2\sqrt{5}}{4} = \frac{3 - \sqrt{5}}{2} = \frac{1 - \sqrt{5} + 2}{2} = \frac{1 - \sqrt{5}}{2} + \frac{2}{2} = \frac{1 - \sqrt{5}}{2} + 1 = \psi + 1$\
Fib($k + 1$) = $\frac{\phi^{k-1}(\phi^2) - \psi^{k - 1}(\psi^2)}{\sqrt{5}} = \frac{\phi^{k + 1} - \psi^{k + 1}}{\sqrt{5}}$\
Therefore, Fib($k + 1$) = $\frac{\phi^{k + 1} - \psi^{k + 1}}{\sqrt{5}}$. It follows by induction that Fib($n$) = $(\phi^n - \psi^n)/\sqrt{5}$ for every $n \in \mathbb{N}$.

Now, we will show that Fib($n$) is the closest integer to $\phi^n/\sqrt{5}$.

As $\lim_{n \to \infty} \psi^n = 0$, so $\lim_{n \to \infty} \frac{\phi^n}{\sqrt{5}} = \lim_{n \to \infty} \frac{\phi^n - \psi^n}{\sqrt{5}}$. This is because $\psi \approx -0.618$ and as it gets exponentiated to higher power, it approaches zero.
```{r psi}
psi <- (1 - sqrt(5)) / 2
plot(x = 1:25, y = sapply(1:25, function(x) psi^x))
```

The proof can also be visualized here, where the difference between approximation and Fibonacci number is displayed on the y-axis.
```{r comparison}
phi <- (1 + sqrt(5)) / 2
fib <- function(x) {
  (phi^x - psi^x) / sqrt(5)
}
approx <- function(x) {
  phi^x / sqrt(5)
}

error <- abs(sapply(1:100, fib) - sapply(1:100, approx))

plot(x = 1:100, y = error, type = "l")
```

As seen on the graph, the largest error at $n = 0$ is less than 0.5, and for all $n \in \mathbb{N}$, $\phi^n/\sqrt{5}$ is the closest integer to Fib($n$).
```{r largest error}
fib(0)
approx(0)
```
